The tasks to solve for the Language Workbench Competition is relatively trivial.
But it has been designed to illustrate as many properties of language
workbenches as possible. As such, it can serve as a good introduction to the
respective tools.

This tutorial expects that you are somehow familiar with Java and Eclipse and
have heard about \url{EMF} and how it works in general before. We start almost at the
beginning, but not quite :-)

For this tutorial we will use Xtext 2.0, which is at the moment of writing not
finalized, but in its latest stage. Xtext 2.0 will be released with Eclipse
Indigo in June
2011\footnote{\url{http://wiki.eclipse.org/Indigo/Simultaneous_Release_Plan}}.
All tasks could also be solved with Xtext 1.0, but for some of the advanced
tasks Xtext 2.0 will have better support and different approaches. All examples
here will already work with the current builds of Xtext 2.0 and
release will be soon after the first publication of LWC11.

\subsection{Task Description}
The LWC11 task is divided into the following phases, and the structure of this
document follows them:

\subsubsection{Phase 0 - Basics}

\begin{compactitem}
    \item 0.1 Simple (structural) DSL without any fancy expression language or
    such. Build a simple data definition language to define entities with
    properties. Properties have a name and a type. It should be possible to use
    primitive types for properties, as well as other Entities.
    \item 0.2 Code generation to GPL such as Java, C\#, C++ or XML. Generate
    Java Beans (or some equivalent data structure in C\#, Scala, etc.) with
    setters, getters and fields for the properties.
    \item 0.3 Simple constraint checks such as name-uniqueness For example,
    check the name uniqueness of the properties in the entities.
    \item 0.4 Show how to break down a (large) model into several parts, while
    still cross-referencing between the parts.
\end{compactitem}

\subsubsection{Phase 1 - Advanced}

This phase demonstrates advanced features not necessarily available to the same
extent in every LWB.

\begin{compactitem}
    \item 1.1 Show the integration of several languages. Define a second
    language to define instances of the entities, including assignment of values
    to the properties. This should ideally really be a second language that
    integrates with the first one, not just ``more syntax'' in the same grammar.
    We want to showcase language modularity and composition here.
    \item 1.2 Demonstrate how to implement runtime type systems. The
    initialization values in the instance-DSL must be of the same type as the
    types of the properties.
    \item 1.3 Show how to do a model-to-model transformation. Define an ER-meta
    model (Database, Table, Column) and transform the entity model into an
    instance of this ER meta model.
    \item 1.4 Some kind of visibility/namespaces/scoping for references.
    Integrate namespaces/packages into the entity DSL and make sure in the
    instance DSL you can only assign values to the properties of the respective
    entity.
    \item 1.5 Integrating manually written code (again in Java, C\# or C++).
    Integrate derived attributes to entities. Note that if you want, you can
    also define or reuse an expression language that allows defining the
    algorithm for calculating the age directly in the model. Ideally, you will
    show both (manually written 3GL code as well as an expression language).
    \item 1.6 Multiple generators. Generate some kind of XML structure from the
    entity model.
\end{compactitem}

\subsubsection{Phase 2 - Non-Functional}

Phase 2 is intended to show a couple of non-functional properties of the LWB.
The task outlined below does not elaborate on how to do this.

\begin{compactitem}
    \item 2.1 How to evolve the DSL without breaking existing models
    \item 2.2 How to work with the models efficiently in the team
    \item 2.3 Demonstrate Scalability of the tools
\end{compactitem}

\subsubsection{Phase 3 - Freestyle}

Every LWB has its own special ``cool features''. In phase three we want the
participants to show off these features. Please make sure, though, that the
features are built on top of the task described below, if possible.
